
```{r set-options, echo=FALSE, cache=FALSE}
options(width=80)
opts_chunk$set(comment = "##", warning = FALSE, message = FALSE, echo = TRUE, tidy = FALSE, size="small")
read_chunk("../scripts/concat.split.R")
read_chunk("../scripts/df.sorter.R")
read_chunk("../scripts/multi.freq.table.R")
read_chunk("../scripts/random.names.R")
read_chunk("../scripts/row.extractor.R")
read_chunk("../scripts/sample.size.R")
read_chunk("../scripts/stringseed.sampling.R")
read_chunk("../scripts/snippets.R")
```

\part{The Functions}

# Where to Get the Functions

The most current source code for the functions described in this document follow.

To load the functions, you can directly source them from the 2657 R Functions page at github: [https://github.com/mrdwab/2657-R-Functions](https://github.com/mrdwab/2657-R-Functions)

You should be able to load the functions using the following (replace `-----------` with the function name^[The "snippets" in Part III of this document can all be loaded from the script `snippets.R`.]):

```{r rcurl_load_scripts, eval=FALSE}
require(RCurl)
baseURL = c("https://raw.github.com/mrdwab/2657-R-Functions/master/")
source(textConnection(getURL(paste0(baseURL, "scripts/-----------.R"))))
```

\cleardoublepage

# concat.split

```{r concatsplit}
```

\cleardoublepage

# df.sorter

```{r dfsorter}
```

\cleardoublepage

# multi.freq.table

```{r multifreqtable}
```

\cleardoublepage

# RandomNames

```{r randomnames}
```

\cleardoublepage

# row.extractor

```{r rowextractor}
```

\cleardoublepage

# sample.size

```{r samplesize}
```

\cleardoublepage

# stringseed.sampling

```{r stringseedsampling}
```

\cleardoublepage
