
```{r set-options, echo=FALSE, cache=FALSE}
options(width=80)
opts_chunk$set(comment = "", warning = FALSE, message = FALSE, echo = TRUE, tidy = FALSE, size="small")
read_chunk("../scripts/df.sorter.R")
```

# df.sorter

The `df.sorter` function allows you to sort a `data.frame` by columns or rows or both. You can also quickly subset data columns by using the `var.order` argument.

## Arguments

* `data`: the source `data.frame`.
* `var.order`: the new order in which you want the variables to appear.
    * Defaults to `names(data)`, which keeps the variables in the original order.
    * Variables can be referred to either by a vector of their index numbers or by a vector of the variable name; partial name matching also works, but requires that the partial match identifies similar columns uniquely (see examples).
    * Basic subsetting can also be done using `var.order` simply by omitting the variables you want to drop.
* `col.sort`: the columns *within* which there is data that need to be sorted.
    * Defaults to `NULL`, which means no sorting takes place.
    * Variables can be referred to either by a vector of their index numbers or by a vector of the variable names; full names must be provided.
* `at.start`: Should the pattern matching be from the start of the variable name? Defaults to "TRUE".

> NOTE: If you are sorting both by variables and within the columns, the `col.sort` order should be based on the location of the columns in the *new* `data.frame`, not the original `data.frame`.

## Examples

```{r df_sorter_examples}
# Load the function!
require(RCurl)
baseURL = c("https://raw.github.com/mrdwab/2657-R-Functions/master/")
source(textConnection(getURL(paste0(baseURL, "scripts/df.sorter.R"))))

# Make up some data
set.seed(1)
dat = data.frame(id = rep(1:5, each=3), times = rep(1:3, 5),
                 measure1 = rnorm(15), score1 = sample(300, 15),
                 code1 = replicate(15, paste(sample(LETTERS[1:5], 3), 
                                             sep="", collapse="")),
                 measure2 = rnorm(15), score2 = sample(150:300, 15), 
                 code2 = replicate(15, paste(sample(LETTERS[1:5], 3), 
                                             sep="", collapse="")))
# Preview your data
dat
# Change the variable order, grouping related columns
# Note that you do not need to specify full variable names,
#    just enough that the variables can be uniquely identified
head(df.sorter(dat, var.order = c("id", "ti", "cod", "mea", "sco")))
# Same output, but with a more awkward syntax
head(df.sorter(dat, var.order = c(1, 2, 5, 8, 3, 6, 4, 7)))
# As above, but sorted by 'times' and then 'id'
head(df.sorter(dat, var.order = c("id", "tim", "cod", "mea", "sco"), 
               col.sort = c(2, 1)))
# Drop 'measure1' and 'measure2', sort by 'times', and 'score1'
head(df.sorter(dat, var.order = c("id", "tim", "sco", "cod"), 
               col.sort = c(2, 3)))
# As above, but using names
head(df.sorter(dat, var.order = c("id", "tim", "sco", "cod"), 
               col.sort = c("times", "score1")))
# Just sort by columns, first by 'times' then by 'id'
head(df.sorter(dat, col.sort = c("times", "id")))
head(df.sorter(dat, col.sort = c("code1"))) # Sorting by character values
# Pattern matching anywhere in the variable name
head(df.sorter(dat, var.order= "co", at.start=FALSE))
```

## To Do

* Add an option to sort ascending or descending---at the moment, not supported.

\newpage
