
```{r set-options, echo=FALSE, cache=FALSE}
options(width=80)
opts_chunk$set(comment = "##", warning = FALSE, message = FALSE, echo = TRUE, tidy = FALSE, size="small")
read_chunk("../scripts/concat.split.R")
read_chunk("../scripts/snippets.R")
```

# concat.split

The `concat.split` function takes a column with multiple values, splits the values into a list or into separate columns, and returns a new `data.frame`.

## Arguments

* `data`: the source `data.frame`.
* `split.col`: the variable that needs to be split; can be specified either by the column number or the variable name.
* `sep`: the character separating each value (defaults to `","`). 

    > ***Note***: *If using `structure = "compact"`, the value for `sep` can only be a single character. See the "Advanced Usage" example of how to specify multiple characters for batch conversion of columns.*

* `structure`: Can be either `"compact"`, `"expanded"`, or `"list"`. Defaults to `"compact"`.
    * `"compact"` creates as many columns as the maximum *length* of the resulting split. This is the most useful general-case application of this function.
    * When the input is numeric, `"expanded"` creates as many columns as the maximum *value* of the input data. This is most useful when converting to `mode = "binary"`.
    * `"list"` creates a single new column that is structurally a `list` within a `data.frame`.
* `mode`: can be either `binary` or `value` (where `binary` is default and it recodes values to `1` or `NA`, like Boolean data, but without assuming `0` when data is not available). This setting only applies when `structure = "expanded"`; an warning message will be issued if used with other structures.
* `drop.col`: logical (whether to remove the original variable from the output or not; defaults to `TRUE`).
* `fixed`: Is the input for the `sep` value *fixed*, or a *regular expression*? When `structure = "expanded"` or `structure = "list"`, it is possible to supply a a regular expression containing the characters to split on. For example, to split on `","`, `";"`, or `"|"`, you can set `sep = ",|;|\\|"` or `sep = "[,;|]"`, and `fixed = FALSE` to split on *any* of those characters. 

## Examples

First load some data from a CSV stored at [github](http://github.com). The URL is an HTTPS, so we need to use `getURL` from `RCurl`.

```{r concat_split_data_loading}
require(RCurl)
baseURL = c("https://raw.github.com/mrdwab/2657-R-Functions/master/")
temp = getURL(paste0(baseURL, "data/concatenated-cells.csv"))
concat.test = read.csv(textConnection(temp))
rm(temp)

# How big is the dataset?
dim(concat.test)
# Just show me the first few rows
head(concat.test)
```

Notice that the data have been entered in a very silly manner. Let's split it up!

```{r concat_split_examples, warning=TRUE}
# Load the function!
# require(RCurl)
# baseURL = c("https://raw.github.com/mrdwab/2657-R-Functions/master/")
source(textConnection(getURL(paste0(baseURL, "scripts/concat.split.R"))))

# Split up the second column, selecting by column number
head(concat.split(concat.test, 2))
# ... or by name, and drop the offensive first column
head(concat.split(concat.test, "Likes", drop.col = TRUE))
# The "Hates" column uses a different separator:
head(concat.split(concat.test, "Hates", sep = ";", drop.col = TRUE))
# You'll get a warning here, when trying to retain the original values
head(concat.split(concat.test, 2, mode = "value", drop.col = TRUE))
# Try again. Notice the differing number of resulting columns
head(concat.split(concat.test, 2, structure = "expanded", 
                  mode = "value", drop.col = TRUE))
# Let's try splitting some strings... Same syntax
head(concat.split(concat.test, 3, drop.col = TRUE))
# Split up the "Likes column" into a list variable; retain original column
head(concat.split(concat.test, 2, structure = "list", drop.col=FALSE))
# View the structure of the output for the first 10 rows to verify 
# that the new column is a list; note the difference between "Likes"
# and "Likes_list".
str(concat.split(concat.test, 2, structure = "list", 
                 drop.col=FALSE)[1:10, c(2, 5)])
```

## Advanced Usage

It is also possible to use `concat.split` to split multiple columns at once. This can be done in stages, or it can be all wrapped in nested statements, as follows:

```{r concat_split_advanced, eval=FALSE}
do.call(cbind, 
        c(concat.test[1], 
          lapply(1:(ncol(concat.test)-1), 
                 function(x) {
                   splitchars = c(",", ",", ";")
                   concat.split(concat.test[-1][x], 1,
                                splitchars[x],
                                drop.col=TRUE)
                 })))
```

In the example above:

* The `lapply()` function is applied to each column in the `data.frame` except the first one.
* Before applying the `concat.split` function, we enter a vector of the characters on which we should split, in the same order as the columns. Here, the first two columns are separated by commas, and the third is separated by a semicolon.
* The `concat.split` function arguments can then be included as you would if splitting a single column.
* We use `do.call(cbind, ...)` to "bind" the data together by columns. Since we had dropped the first column for the `lapply` step, we add that back in at this stage.

```{r concat_split_advanced_2}
# Show just the first few lines, compact structure
# Note that the split characters must be specified
#   in the same order that lapply will encounter them
head(do.call(cbind, 
        c(concat.test[1], 
          lapply(1:(ncol(concat.test)-1), 
                 function(x) {
                   splitchars = c(",", ",", ";")
                   concat.split(concat.test[-1][x], 1,
                                splitchars[x],
                                drop.col=TRUE)
                 }))))

# Show just the first few lines, Boolean mode
# Note the use of a regular expression for sep
#   and the setting of fixed to FALSE
head(do.call(cbind, 
        c(concat.test[1], 
          lapply(1:(ncol(concat.test)-1), 
                 function(x) {
                   concat.split(concat.test[-1][x], 1,
                                sep = "[,;]",
                                structure = "expanded",
                                fixed = FALSE,
                                drop.col=TRUE)
                 }))))


# Show just the first few lines, value mode
head(do.call(cbind, 
        c(concat.test[1], 
          lapply(1:(ncol(concat.test)-1), 
                 function(x) {
                   concat.split(concat.test[-1][x], 1,
                                sep = "[,;]",
                                structure = "expanded",
                                mode = "value",
                                fixed = FALSE,
                                drop.col=TRUE)
                 }))))

# Show just the first few lines, list output mode
head(do.call(cbind, 
        c(concat.test[1], 
          lapply(1:(ncol(concat.test)-1), 
                 function(x) {
                   concat.split(concat.test[-1][x], 1,
                                sep = "[,;]",
                                structure = "list",
                                fixed = FALSE,
                                drop.col=TRUE)
                 }))))
```

## References

See: [http://stackoverflow.com/q/10100887/1270695](http://stackoverflow.com/q/10100887/1270695)    
The `"condensed"` setting was inspired by an answer from David Winsemius (@DWin) to a question at Stack Overflow. See: [http://stackoverflow.com/a/13924245/1270695](http://stackoverflow.com/a/13924245/1270695) 

\cleardoublepage
