
```{r set-options, echo=FALSE, cache=FALSE}
options(width=80)
opts_chunk$set(comment = "##", warning = FALSE, message = FALSE, echo = TRUE, tidy = FALSE, size="small")
read_chunk("../scripts/table2df.R")
```

# table2df

The `table2df` function takes an object of class `table`, `ftable`, and `xtabs` and converts them to `data.frame`s, retaining as many of the name details as possible.

## Arguments

* `mytable`: The table object you want to convert into a `data.frame`. This can be a saved object or you can make the call to `table`, `ftable`, or `xtabs` as the `mytable` argument to this function.
* `as.multitable`: Logical; defaults to `FALSE`. Some methods, for instance `xtabs` and `table`, will create an array of tables as the output when more than two variables are being tabulated. 
    * If `as.multitable` is `TRUE`, the function will return a list of `data.frame`s. 
    * If `as.multitable` is `FALSE`, the function will convert the object to an `ftable` object before performing the transformation.
* `direction`: Can be either `"long"` or `"wide"`. 
    * If `"long"`, the frequencies will all be tabulated into a single column. This is the same behavior you will generally get if you used `as.data.frame()` on a `table` object.
    * If `"wide"`, the tabular format is retained.

## Examples

First, we will create a dataset containing information about a person's gender; whether they are left handed, right handed, or ambidextrous; their favorite colors and favorite shapes; and the operating system on their computer.

```{r table2df_data}
set.seed(1)
handedness <- data.frame(
  gender = sample(c("Female", "Male", "Unknown"), 200, replace = TRUE),
  handedness = sample(c("Right", "Left", "Ambidextrous"), 
                      200, replace = TRUE, prob = c(.7, .2, .1)),
  fav.col = sample(c("Red", "Orange", "Yellow", "Green", "Blue", 
                     "Indigo", "Violet", "Black", "White"), 
                   200, replace = TRUE),
  fav.shape = sample(c("Triangle", "Circle", "Square", "Pentagon", "Hexagon",
                       "Oval", "Octagon", "Rhombus", "Trapezoid"),
                     200, replace = TRUE),
  computer = sample(c("Win", "Mac", "Lin"), 200, replace = TRUE,
                    prob = c(.5, .25, .25)))

# Preview the data
list(head(handedness), tail(handedness))
```
```{r table2df_examples}
# Load the function!
require(RCurl)
baseURL = c("https://raw.github.com/mrdwab/2657-R-Functions/master/")
source(textConnection(getURL(paste0(baseURL, "scripts/table2df.R"))))

# A very basic table
HT1 <- with(handedness, table(gender, handedness))
HT1
table2df(HT1)
table2df(HT1, direction = "long")

# Another basic table
HT2 <- with(handedness, table(fav.col, computer))
HT2
table2df(HT2)

# This will create multiple tables, one for each possible computer value
HT3 <- with(handedness, table(gender, fav.col, computer))
HT3
# Default settings
table2df(HT3)
# As a list of data.frames
table2df(HT3, as.multitable = TRUE)
# As above, but with the output in long format
#   Only showing the first three lines of each data.frame
lapply(table2df(HT3, as.multitable = TRUE, direction = "long"), head, 3)

# Applied to an ftable
HT4 <- ftable(handedness, 
              col.vars="fav.col", 
              row.vars=c("gender", "computer"))
HT4
table2df(HT4)
```

### Other examples

Here are a few other examples to try using some of the inbuilt datasets available in R.

```{r eval=FALSE}
table2df(xtabs(cbind(ncases, ncontrols) ~ ., data = esoph))
table2df(xtabs(cbind(ncases, ncontrols) ~ ., data = esoph), 
         direction = "long")
table2df(xtabs(cbind(ncases, ncontrols) ~ ., data = esoph), 
         as.multitable = TRUE, direction = "long")
table2df(xtabs(cbind(ncases, ncontrols) ~ ., data = esoph), 
         as.multitable = TRUE, direction = "wide")
table2df(ftable(Sex ~ Class + Age, data = x))
```

## References

The `expand.grid` method for remaking the columns from an `ftable` was described by Kohske at [http://stackoverflow.com/a/6463137/1270695](http://stackoverflow.com/a/6463137/1270695).
