
```{r set-options, echo=FALSE, cache=FALSE}
options(width=80)
opts_chunk$set(comment = "##", warning = FALSE, message = FALSE, echo = TRUE, tidy = FALSE, size="small")
read_chunk("../scripts/row.extractor.R")
```

# row.extractor

The `row.extractor` function takes a `data.frame` and extracts rows with the `min`, `median`, or `max` values of a given variable, or extracts rows with specific quantiles of a given variable.

## Arguments

* `data`: the source `data.frame`.
* `extract.by`: the column which will be used as the reference for extraction; can be specified either by the column number or the variable name.
* `what`: options are `min` (for all rows matching the minimum value), `median` (for the median row or rows), `max` (for all rows matching the maximum value), or `all` (for `min`, `median`, and `max`); alternatively, a numeric vector can be specified with the desired quantiles, for instance `c(0, .25, .5, .75, 1)`

## Examples

```{r row_extractor_examples}
# Load the function!
require(RCurl)
baseURL = c("https://raw.github.com/mrdwab/2657-R-Functions/master/")
source(textConnection(getURL(paste0(baseURL, "scripts/row.extractor.R"))))

# Make up some data
set.seed(1)
dat = data.frame(V1 = 1:50, V2 = rnorm(50), 
                 V3 = round(abs(rnorm(50)), digits=2), 
                 V4 = sample(1:30, 50, replace=TRUE))
# Get a sumary of the data
summary(dat)
# Get the rows corresponding to the 'min', 'median', and 'max' of 'V4'
row.extractor(dat, 4) 
# Get the 'min' rows only, referenced by the variable name
row.extractor(dat, "V4", "min") 
# Get the 'median' rows only. Notice that there are two rows 
#    since we have an even number of cases and true median 
#    is the mean of the two central sorted values
row.extractor(dat, "V4", "median") 
# Get the rows corresponding to the deciles of 'V3'
row.extractor(dat, "V3", seq(0.1, 1, 0.1)) 
```

## To Do

* Add some error checking to make sure a valid `what` is provided.

## References

`which.quantile` function by [cbeleites](http://stackoverflow.com/users/755257/cbeleites)  
See: [http://stackoverflow.com/q/10256503/1270695](http://stackoverflow.com/q/10256503/1270695)

\cleardoublepage
