
```{r set-options, echo=FALSE, cache=FALSE}
options(width=80)
opts_chunk$set(comment = "", warning = FALSE, message = FALSE, echo = TRUE, tidy = FALSE, size="small")
read_chunk("../scripts/sample.size.R")
```

# sample.size

The `sample.size` function either calculates the optimum survey sample size when provided with a population size, or the confidence interval of using a certain sample size with a given population. It can be used to generate tables (`data.frame`s) of different combinations of inputs of the following arguments, which can be useful for showing the effect of each of these in sample size calculation.

## The Arguments

* `population`: The population size for which a sample size needs to be calculated.
* `samp.size`: The sample size. 
    * This argument is only used when calculating the confidence interval, and defaults to `NULL`.
* `c.lev`: The desired confidence level. Defaults to a reasonable 95%.
* `c.int`: The confidence interval.
    * This argument is only used when calculating the sample size.
    * If not specified when calculating the sample size, defaults to 5% and a message is provided indicating this; this is also the default action if `c.int = NULL`.
* `what`: Should the function calculate the desired sample size or the confidence interval?
    * Accepted values are `"sample"` and `"confidence"` (quoted), and defaults to "`sample`".
* `distribution`: Response distribution. Defaults to 50%, which will give you the largest sample size. 

## Examples

```{r sample_size_exampes}
# Load the function!
require(RCurl)
baseURL = c("https://raw.github.com/mrdwab/2657-R-Functions/master/")
source(textConnection(getURL(paste0(baseURL, "scripts/sample.size.R"))))
# What should our sample size be for a population of 300?
# All defaults accepted.
sample.size(population = 300)
# What sample should we take for a population of 300
#   at a confidence level of 97%?
sample.size(population = 300, c.lev = 97)
# What about if we change our confidence interval?
sample.size(population = 300, c.int = 2.5, what = "sample")
# What about if we want to determine the confidence interval
#   of a sample of 140 from a population of 300? A confidence
#   level of 95% is assumed.
sample.size(population = 300, samp.size = 140, what = "confidence")
```

## Advanced Usage

As the function is vectorized, it is possible to easily make tables with multiple scenarios.

```{r sample_size_advanced}
# What should the sample be for populations of 300 to 500 by 50?
sample.size(population=c(300, 350, 400, 450, 500))
# How does varying confidence levels or confidence intervals
#   affect the sample size?
sample.size(population=300, 
            c.lev=rep(c(95, 96, 97, 98, 99), times = 3),
            c.int=rep(c(2.5, 5, 10), each=5))
# What is are the confidence intervals for a sample of
#   150, 160, and 170 from a population of 300?
sample.size(population=300, 
            samp.size = c(150, 160, 170), 
            what="confidence")
```

Note that the use of `rep()` is required in constructing the arguments for the advanced usage examples where more than one argument takes on multiple values.

## References

See the *2657 Productions News* site for how this function progressively developed^[[http://news.mrdwab.com/2010/09/10/a-sample-size-calculator-function-for-r/](http://news.mrdwab.com/2010/09/10/a-sample-size-calculator-function-for-r/)]. The `sample.size` function is based on the following formulas^[See: Creative Research Systems. (n.d.). *Sample size formulas for our sample size calculator*. Retrieved from: [http://www.surveysystem.com/sample-size-formula.htm](http://www.surveysystem.com/sample-size-formula.htm). Archived on 07 August 2012 at [http://www.webcitation.org/69kNjMuKe](http://www.webcitation.org/69kNjMuKe).]:

$$
\large 
\begin{array}{rcl}
ss &=& \frac{-Z^2\times p\times(1-p)}{c^2} \\ \\
pss &=& \frac{ss}{1+\frac{ss-1}{pop}}
\end{array}
$$

\newpage
