
```{r set-options, echo=FALSE, cache=FALSE}
options(width=80)
opts_chunk$set(comment = "##", warning = FALSE, message = FALSE, echo = TRUE, tidy = FALSE, size="small")
read_chunk("../scripts/concat.split.R")
read_chunk("../scripts/snippets.R")
```

# stratified

The `stratified` function samples from a `data.frame` in which one of the columns represents an "identifier" variable and one of the columns represents a "stratification" or "grouping" variable. The result is a new `data.frame` with the specified number of samples from each group.

## Arguments

* `df`: The source `data.frame`.
* `id`: Your "ID" variable.
* `group`: Your grouping variable.
* `size`: The desired sample size. If `size` is a decimal less than 1, a proportionate sample would be drawn. If `size` is a positive integer, it will be assumed that you want the same number of samples from each group.

    > ***Note***: Because of how computers deal with floating-point arithmetic, and because R uses a "round to even" approach, the `size` per strata that results when specifying a proportionate sample may be slightly higher or lower per strata than you might have expected.

* `seed`: The seed that you want to use (using `set.seed()`), if any. Defaults to `NULL`. 

    > ***Note***: This is different from using `set.seed()` before using the function. Setting a seed using this argument is equivalent to using `set.seed(seed)` each time that you go to take a sample from a different group (in other words, the same seed is used for each group).  
    
* `...`: Further arguments to be passed to the `sample()` function.

## Examples

First, let's make up some data. In the dataset below, we can treat variables "A" and "D" as potential grouping variables.

```{r}
# Generate some sample data to play with
set.seed(1)
dat = data.frame(ID = 1:100, 
                 A = sample(c("AA", "BB", "CC", "DD", "EE"), 100, replace=T),
                 B = rnorm(100), C = abs(round(rnorm(100), digits=1)),
                 D = sample(c("CA", "NY", "TX"), 100, replace=T))

# What do the data look like in general?
summary(dat)
```

Now, let's try different settings applying the `stratified` function.

```{r}
# Load the function!
require(RCurl)
baseURL = c("https://raw.github.com/mrdwab/2657-R-Functions/master/")
source(textConnection(getURL(paste0(baseURL, "scripts/stratified.R"))))

# Let's take a 10% sample from all -A- groups, seed = 1
stratified(dat, "ID", "A", .1, seed = 1)

# Let's take 5 samples from all -D- groups, seed = 1, specified by column number
stratified(dat, 1, 5, 5, 1)
```

## To Do

* Error handling for when fixed sample size is specified but some strata have fewer observations than specified sample size.
* Option to specify vector of required samples
* set.seed examples

## References

The evolution of this function can be found at the following URLs: 

1. [http://news.mrdwab.com/2011/05/15/stratified-random-sampling-in-r-beta/](http://news.mrdwab.com/2011/05/15/stratified-random-sampling-in-r-beta/)
1. [http://news.mrdwab.com/2011/05/20/stratified-random-sampling-in-r-from-a-data-frame/](http://news.mrdwab.com/2011/05/20/stratified-random-sampling-in-r-from-a-data-frame/)
1. [http://stackoverflow.com/a/9714207/1270695](http://stackoverflow.com/a/9714207/1270695)

The version here is entirely reworked and does not require an additional package to be loaded.

\cleardoublepage